Tempus: Probabilistic Network Latency Verifier

### 2023-02-01 paper/1_intro.tex
[Paper] Changed Intro

\section{Introduction}
Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
routing information so that a packet can reach their intended destination efficiently, even in the 
event of a component failure.
Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
It is hard for a network engineer to reason about whether their configured network fulfilled some 
intended property in various possible states of the network. 
Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
to help them in this task, namely, network control plane verifier.

Popular control plane verifiers, like ARC \cite{arc}, are formulated to answer 
questions about a given property deterministically. 
In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
expected to give a yes or no answer about the property (e.g. two nodes are reachable).
While useful to some extent, this kind of models are sometimes too restrictive since network 
operators are usually able to tolerate small fraction of failure. 
For example, a given network might have an availability SLA of 99.999\%. 
This kind of probabilistic properties have been the focus of a more recent works like NetDice 
\cite{steffen2020probabilistic}.

The network property itself can also be divided into two kinds. 
Quantitative properties like reachability and loop existence, are the common properties that are 
studied by most of the existing verifiers. 
More recent work, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
properties like link bandwidth violation for a given traffic.

In this work, we're exploring the other side of the network performance metric: latency. 
Certain network deployment often necessitates some latency requirement such as an ISP that has 
latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
We proposed a verification framework to probabilistically verify the latency property of packets 
traversing from a source to destination node under various failure scenarios, by using latency 
information of each components in the network.

>>>>

\section{Introduction}
The intricate configuraton of network devices is notoriously hard to get right, forcing network 
engineers to accept the reality of Murphy's Law or getting some help from a tool to ensure their 
correctness, namely, network verifier.
Over the last decade, there has been a lot of development in the area of network verification 
generally \cite{hsa}\cite{veriflow}.
While their approach varies a lot (e.g. data plane vs. control plane verification, deterministic vs. 
probabilistic), most of these verifiers shares a common focus, primarily fixated on verifying 
\textit{qualitative} properties -- the convergent functional behavior of a network under various failure 
scenarios -- such as reachability and loop detection.

However, network engineers also have the need to verify \textit{quantitative} properties, such as properties 
related to bandwidth and latency.
Still, barring some recent works about the verification of link load violation \cite{qarc}, work in this area 
remain sparse, despite its undoubtable importance.

In this work, we're exploring the verification of the other side of the network performance metric: latency. 
Certain network deployment often necessitates some latency requirement such as an ISP that has 
latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
We proposed a verification framework to probabilistically verify the latency property of packets 
traversing from a source to destination node under various failure scenarios, by using latency 
measurements of the components in the network.

We introduce the design and implementation of \tool, a probabilistic network latency verification 
framework.
\tool will use the latency information from relevant component measurements (e.g. router queue length) 
to infer the latency distribution of said component.
Assuming that future traffic is i.i.d., we then employ a numerical convolution method to combine the
latency distribution of each relevant components together to produce the end-to-end latency distribution 
of an src-dst router pair.

To determine said relevant components, we built \tool on top of an existing qualitative verification 
framework to efficiently explore the path used for an src-dst router pair given various scenarios of 
failures.
By obtaining the end-to-end latency distribution from this information and the latency measurement, 
we could analyze the statistical properties of the src-dst pair latency distribution to probabilistically 
argue about the latency properties.

We also introduced two optimization techniques on top of this framework to reduce the verification time 
by multiple orders of magnitude.
These two optimization techniques rely on the symmetry that we have found in the quantitative verifier 
when brought into the context of latency verification.

Our evaluation shows that the verification framework, with the help of our optimization techniques, 
could accomplish the verification task on various WAN and datacenter topologies in the order of 
minutes?

With this work, we make the following technical contributions:
\begin{itemize}
    \item \textbf{Novel temporal verification framework} using numerical convolution of network 
        components' latency measurement.
    \item \textbf{Two optimization techniques} in said verification framework by exploiting 
        symmetry in the failure exploration.
    \item \textbf{\tool}, an implementation of our verification framework and its optimization 
        on Julia.
\end{itemize}

The code for this work is open-sourced at link. %TODO: change link
This work does not raise any ethical issues.

### 2023-02-01 paper/3_functional.tex
[Paper] Changed Intro

\section{Functional Verification}
For functional property verification, NetDice \cite{steffen2020probabilistic} 
has laid the way for verifying reachability between two nodes under failure in a 
probabilistic and efficient manner.
To contextualize our modification, we will briefly describe some parts of 
NetDice's network encoding and exploration algorithm as a prelude to our alteration.
For a more complete complete description of NetDice, please refer to the original 
paper.

>>>>

\section{Functional Verification}
For functional property verification, NetDice \cite{netdice} 
has laid the way for verifying reachability between two nodes under failure in a 
probabilistic and efficient manner.
To contextualize our modification, we will briefly describe some parts of 
NetDice's network encoding and exploration algorithm as a prelude to our alteration.
For a more complete complete description of NetDice, please refer to the original 
paper.

### 2023-02-01 paper/6_related.tex
[Paper] Changed Intro

\textbf{Probabilistic Verifier}
More recent works have tried to address this very issue.
Works like ProbNetKAT \cite{probnetkat} (a probabilistic verifier based on NetKAT 
\cite{netkat}), NetDice \cite{steffen2020probabilistic}, and SRE \cite{sre} provides 
the user ability to define probabilistic failure model, where a given component 
of the network can fail with a given probability.

@@

\textbf{Quantitative Verifier}
Some recent works are an exception to this trend, however.
Works like QARC \cite{subramanian2020detecting} (a quantitative verifier based on 
ARC \cite{arc}) has developed a framework to verify one particular quantitative 
property: excess bandwidth on a certain link.
While QARC is a deterministic verifier, like Tempus, they focuses on a more quantitative 
property. 
Unlike Tempus however, this work is orthogonal to verifying latency and thus complementary 
to Tempus.

>>>>

\textbf{Probabilistic Verifier}
More recent works have tried to address this very issue.
Works like ProbNetKAT \cite{probnetkat} (a probabilistic verifier based on NetKAT 
\cite{netkat}), NetDice \cite{netdice}, and SRE \cite{sre} provides 
the user ability to define probabilistic failure model, where a given component 
of the network can fail with a given probability.

@@

\textbf{Quantitative Verifier}
Some recent works are an exception to this trend, however.
Works like QARC \cite{qarc} (a quantitative verifier based on 
ARC \cite{arc}) has developed a framework to verify one particular quantitative 
property: excess bandwidth on a certain link.
While QARC is a deterministic verifier, like Tempus, they focuses on a more quantitative 
property. 
Unlike Tempus however, this work is orthogonal to verifying latency and thus complementary 
to Tempus.
